Word Learning as Network Growth: a Cross-linguistic Analysis

### 2018-01-23 cogsci/cogsci.tex
cogsci up-date and packrat

\title{The Role of Early Semantic and Phonological Networks in Noun Learning: a
Cross-linguistic Analysis}

@@

\begin{abstract}
Children tend to learn first the words that have a high density
neighborhood in the lexical network (i.e., words that are connected to a
variety of other words). However, the reason why connectivity influences
learning is still not very well understood. Previous studies suggested
this effect could be the consequence of the underlying learning
mechanism. However, these studies focused on semantic networks and
English data only, thus severely limiting the generality of the
findings. Here we provide a systematic analysis of network growth models
in both the phonological and the the semantic domains, and across 8
languages. We replicate the results obtained by Hills et al. (2009) with
English semantic networks, show that the same mechanism drives the
growth of both semantic and phonological networks, and demonstrate that
these effects hold cross-linguistically.

\textbf{Keywords:}
keywords: semantic network, phonological network, word learning,
semantic
\end{abstract}

@@

A crucial question in language development is to understand why children
learn some words at an earlier age than others. Previous studies
documented various word properties that influence the age of
acquisition. For example, it was shown that, within a lexical category,
words that are more frequent in child-directed speech are acquired
earlier (Goodman, Dale, \& Li, 2008). Other predictors include word
length, the mean length of utterances in which the word occurs, and how
concrete the word is (see Braginsky, Yurovsky, Marchman, \& Frank,
2016). Besides these word-level properties, researchers found that the
lexical structure (that is, the network that specifies how words relate
to each other) also influences the learning process. For instance,
children tend to produce first the words that have higher neighborhood
density (i.g., high connectivity in the network) both at the
phonological and the semantic level (Stokes, 2010; Storkel, 2009).
However, it is still not very well understood \emph{why} connectivity in
the lexical structure influences learning. Steyvers \& Tenenbaum (2005)
suggested that this effect is nothing but the consequence of how these
structures get constructed in the child's mind. According to this
explanation (known as Preferential Attachment), highly connected words
in the child's lexicon tend to ``attract'' more words over time, in a
rich-get-richer scenario. Nonetheless, Hills, Maouene, Maouene, Sheya,
\& Smith (2009) found that what biases the learning is not the
connectivity in the child's internal lexicon but the connectivity in the
learning environment.

These analyses, however, focused on growth in semantic networks only. A
recent study by Stella, Beckage, \& Brede (2017) analysed both semantic
and phonological connectivity, but assumed only one growth model
(preferential acquisition). More importantly, almost all available
studies are limited in their generality because they focus almost
exlusively on English data. The novely of this study is that it
investigtes, for the first time, whether phonological networks grow by
preferential attachement or preferential acquisition. It also provides a
systematic comparision between network growth scenarios in both the
phonological and the semantic domains and assess their relative
contribution to the learning process. Moreover, it tests the generality
of the findings across eight languages.

The paper is organized as follows. First, we describe the dataset we
used, the procedure we followed to construct both semantic and
phonological networks, and the network growth scenarios we tested. Next,
we show the results of the various analyses we conducted comparing
network growth scenarios across languages. Then we compare the overall
contribution of the networks in the learning process relative to other
known predictors of age of acquisition (frequency and length). Finally
we discuss the major results in relation to known experimental facts in
language development.

@@

We used data from Wordbank (Frank, Braginsky, Yurovsky, \& Marchman,
2017), an open repository aggregating cross-linguistic language
developmental data of the MacArthur-Bates Communicative Development
Inventory (CDI). The latter consists of parent report vocabulary
checklists. We used the \emph{Words and Sentence} version which contains
the productive vocabulary of toddlers (age varied between 16 to 35
months). We included words that were produced by at least 50\% of
children, and we considered that these words were learned at the month
when this percentage was reached (Goodman et al., 2008). Moreover,
following previous studies (Hills et al., 2009; e.g., Storkel, 2009), we
restricted the analysis to the category of nouns, since there are known
discrepancies between the learning of nouns, predicates, and function
words (e.g., Braginsky et al., 2016). We obtained these nouns in eight
languages: Croatian, Danish, English, Italian, Norwegian, Russian,
Spanish, and Turkish. We used the subset of nouns that had entries in
the Florida Association Norms (see below). Since these norms are
available only in English, we mapped words onto translation equivalents
across CDI forms. This allowed us to use the English association norms
across languages. Table \ref{tab:stats} gives an overview of the data
used. The translation of non-English words is still an ongoing process.
Note, however, that all languages have at least 60\% of nouns
translated.

\begin{table}[H]
\centering
\begin{tabular}{rlrrr}
  \hline
 & language & total & translated & normed \\ 
  \hline
1 & Croatian & 253 & 177 & 170 \\ 
  2 & Danish & 295 & 198 & 187 \\ 
  3 & English & 296 & 296 & 274 \\ 
  4 & Italian & 311 & 203 & 194 \\ 
  5 & Norwegian & 305 & 193 & 186 \\ 
  6 & Russian & 311 & 311 & 285 \\ 
  7 & Spanish & 240 & 173 & 163 \\ 
  8 & Turkish & 293 & 175 & 164 \\ 
   \hline
\end{tabular}
\caption{\label{tab:stats}Total number of productive nouns in the CDI (left). We used a subset of these nouns that have available English translations (middle), as well as entries in the Free Association Norms (right).} 
\end{table}

@@

We used as index of semantic relatedness the Florida Free Association
Norms (Nelson, McEvoy, \& Schreiber, 1998). This dataset was collected
by giving adult participants a word (the cue), and asking them to write
the first word that comes to mind (the target). For example, when given
the word ``ball'', they might answer with the word ``game''. A pair of
nodes were connected by a directed link from the cue to the target if
there was a cue-target relationship between these nodes in the
association norms. Each node was characterized by its ``indegree'',
which represents the number of links for which the word was the target.
Following Hills et al. (2009), we used as a proxy for the learning
environment, the network constructed using the full CDI data in each
language (e.g., in English this corresponds to the network by 30
months).

@@

\caption{\label{fig:pred_ind}Evaluation of individual network growth models using both  one-sample t-test over the z-score distribution of learned words (dotted), and model fitting (solid). Each dot represents, in the first case, the mean of the z-score distribution of the learned words with ranges representing 95\% confidence interval, and in the second case, the mean of the posterior distribution of the model's parameter, with ranges reppresenting 95\% credible intervals (computed using the highest density intervals).}\label{fig:pred_ind_img}
\end{figure*}
\end{CodeChunk}

We used as a measure of phonological relatedness between two nodes the
Levenshtein (edit) distance between their phonological forms. The
measure counts the minimum number of operations (insertions, deletions,
substitutions) required to change one string into another. We generated
approximate IPA forms from the orthographic transcription, across
languages, using the open source text-to-speech software
\textbf{\href{http://http://espeak.sourceforge.net/}{Espeak}.}. In
previous studies, two nodes were linked if they had an edit distance of
1 (Stokes, 2010; Storkel, 2009). However, in these previous studies the
network was built using an adult vocabulary. Here, similar to the
approximation we made for the semantic network, we used the full set of
nouns in the CDI data as a proxy for phonological connectivity in the
learning environment. However, since the children's vocabulary contains
very few word pairs with an edit distance of 1, the resulting network
was too sparse and uninformative. Thus, we increased the threshold from
1 to 2, that is, two node were related if their edit distance was equal
to 1 or 2. Each node was characterized with its degree, i.e., the number
of links it shares with other words.

\subsection{Network growth}\label{network-growth}

We tested two network growth scenarios discussed in Hills et al. (2009).
The first one was \emph{Preferential Attachment} (PAT). According to
this mechanism, the network structure of known words predicts what words
will be learned next. A word is more likely to be learned if it is
linked to one of the highly connected nodes in this internal network.
The second mechanism was \emph{Preferential Acquisition} (PAC). In this
model, what predicts what word will enter the lexicon is not
connectivity in the internal network, but connectivity in the learning
environment. Figure \ref{fig:growth} shows an illustration of both
growth scenarios with the same simplified network. For PAT, each
candidate node was characterized with the average degree of the existing
nodes that it would attach to. Thus, according to PAT, the node N1 is
more likely to enter the lexicon first. For PAC, each candidate node was
characterized with its degree in the entire network. According th PAC,
the node N3 is more likely to enter the lexicon first.

\caption{\label{fig:growth}Illustration of the two growth models with the same network. Filled circles (A to D) represent known words, and empty circles (N1 to N3) represent words that have not been learned yet. Back lines represent links that are relevant in each growth scenario, and gray lines represent links that are irrelevant (from Hills et al. 2009).}\label{fig:growth}
\end{figure}
\end{CodeChunk}

\subsection{Connectivity vs.~Age of
acquisition}\label{connectivity-vs.age-of-acquisition}

Preferential attachment and preferential acquisition provide different
generative scenarios for the same end-state. The latter is defined here
as the network corresponding to the oldest age for which we have CDI
data. However in both cases, it is predicted that nodes with the higher
degrees in this end-state are learned earlier than nodes with the lower
degrees. Figure \ref{fig:data_all} shows how the age of acquisition for
each word varies as a function of its degree (or indegree for the
semantic network) in the end-state network. In order to compare this
correlation across languages, the predictor (i.e., the degree) was
centered and scaled. The plot shows a negative correlation between the
month of acquisition and the degree (with the exception of the Russian
semantic network), indicating that nouns with higher degrees are
generally learned earlier.

{\centering \includegraphics{figs/pred_comb_img-1}

\caption{\label{fig:pred_comb_img}Estimates of the growth models' parameters in each language. Dots represent the means of the posteriori distribution of the parameters, and ranges represent 95\% credible intervals (computed using the highest density intervals). }\label{fig:pred_comb_img}
\end{figure*}
\end{CodeChunk}

@@

Following Hills et al. (2009), we evaluated the growth models in two
separate ways. The first evaluation consists in determining, in each
month, the model-dependent growth value distribution of all words that
could possibly be learned at this month, and then computing the z-score
of each learned word with respect to this distribution. For each model,
we tested if the distribution constituted by the z-scores of all learned
words was different from zero, using a one-sample t-test. The results
are shown in Figure \ref{fig:pred_ind}.

The second evaluation consists in fitting an explicit model to the data.
The purpose of this second evaluation is, first, to validate the
findings obtained with the one-sample t-test, and second, to determine
(in a next step) the relative contribution of each network growth
scenario (PAT vs.~PAC) across domains (phonological vs.~semantic).

We proceeded as follows. We calculated the probability that a word
\(w_i\), with a growth value \(d_i\) would enter the lexicon at a given
month, using a softmax function:

@@

Where \(\beta\) is the fitting parameter. A positive value of \(\beta\)
means that words with higher groth values \(d_i\) are acquired first,
and a negative value means that words with lower growth values are
acquired first. The normalization included words that were not yet
learned at the month the word at hand was acquired. We estimated the
parameter \(\beta\) using a Bayesian approach. We started with a uniform
distribution, and at each month, we computed the likelihood function
over words that could possibly enter the lexicon at that month, fit to
the words that have been actually learned at that month (using formula
1). We ended up with a posterior distribution of \(\beta\), which we
summarized in Figure \ref{fig:pred_ind}.

The results from both evaluations were very similar and lead essentially
to the same conclusions. For the semantic networks, the results
replicate Hills et al.'s finding in English, which was that the semantic
network grows by PAC, not by PAT. Moreover, it generalizes this finding
to all other languages (with the exception of Russian). For the
phonological network, the results show that, generally speaking, PAC
fits the developmental trajectory better than PAT in all languages. We
note however that PAT, though weaker, fares better for the phonological
networks than it does for the semantic networks. In the former, it
predicts part of the growth process in some languages such as Croatian,
English, Norwegian and Russian. In the latter, it is rather universally
unpredictive.

Above we evaluated the network growth scenarios individually. As a next
step, we analysed their relative contribution to the learning process.
This was done through adding more fitting parameters to the model, that
is, by substituting \(\beta d_i\) in equation (1) with:
\[\beta_{1} d_{i, 1} + \beta_{2} d_{i, 2} + \beta_{3} d_{i, 3} + \beta_{4} d_{i, 4}\]
where the indices represent the 4 networks: semPAT, semPAC, phonoPAT and
PhonoPAC. Using the same fitting technique, we obtained the values shown
in Figure \ref{fig:pred_comb_img}. In term of growth scenarios, we found
that PAC dominates the learning. In particular, though we found
previously that phonological PAT predicted part of the learning in some
languages when tested individually, here PAT appears to lose its
predictive power when pitted against phonological PAC. In terms of
domain (phonological vs.~semantic), both phonological and semantic
networks appear to contribute to learning, although the phonological
network appears to be stronger and more reliable across languages. In
summary, the findings show that both semantic and phonological networks
grow primarily through preferential acquisition, and that, generally
speaking, semantic and phonological networks both contribute to the
learning process.

@@

\subsection{The big picture: comparison to other known predictors of age
of
acquisition}\label{the-big-picture-comparison-to-other-known-predictors-of-age-of-acquisition}

We saw that the way semantic and phonological information is structured
in the learning environment (i.e., PAC) contribute to noun learning
across languages. However, we know that other factors influence learning
as well (e.g., Braginsky et al., 2016). In what follows, we will
investigate how semantic and phonological connectivity interact with two
other factors. The first one is word frequency, a well studied factor
shown to predict the age of acquisition in a reliable fashion (e.g.
Goodman et al., 2008). The second factor is word length because it
correlates with phonological connectivity (Pisoni, Nusbaum, Luce, \&
Slowiaczek, 1985).

Since PAT was uninformative, we dropped it from this analysis. Thus, we
no longer needed to to fit the growth model month-by-month as in the
previous section. In fact, word utilities in the case of PAC are fixed,
they do not depend on previously learned words. A more direct way to
assess and compare the contribution of PAC in relation to other factors
is through conducting linear regressions, where connectivity in the
learning environment (at both the phonological and semantic level),
frequency and length predict the age of acquisition.

We used the frequency estimates from Braginsky et al. (2016) where
unigram counts were derived based on CHILDES corpora in each language.
For each word, counts included words that shared the same stem (e.g.,
cats counts as cat), or words that were synonymous (e.g.~father counts
as daddy). For word length, we used our generated IPA transcription.
This allowed us to get a more accurate estimate of word length across
languages than the number of characters in the orthographic
transcription.

@@

\caption{\label{fig:regressions_all_img}Estimates of predictor coefficients in the combined model. Ranges indicate 95\% confidence intervals. Fade color indicates estimates of PAC predictors in a model that does not include frequency and length as covariates.}\label{fig:regressions_all_img}
\end{figure}
\end{CodeChunk}

We conducted two analyses. On the one had, we fit a linear regression
for each language, and on the other hand, we fit a linear mixed-effect
to all the data pooled across languages, with language as a random
effect. Figure \ref{fig:regressions_img} shows the coefficient estimate
for each predictor in each language, and figure
\ref{fig:regressions_all_img} shows the coefficient estimates for all
languages combined (all predictors were centered and scaled). The
findings were as follows. Overall, frequency is the most reliable
predictor of age of acquisition. Word length predicts part of learning
in some languages such as Croatian and Norwegian, but not in others
(including English). It remains, however, a significant predictor in the
global model. As for the factors of interest, i.e., semantic and
phonological connectivities, we also found cross-linguistic differences.
The phonological connectivity contributes to learning in languages such
as Croatian, English and Russian, whereas semantic connectivity
contributes to learning in Turkish, Spanish and to some extent in
Croatian, but interestingly not in
English\footnote{Semantic connectivity does not explain variance in English data beyond that explained by frequency and length. This contrasts with the original finding in Hills et al. 2009. However, in this previous studiy, semantic connectivity was not tested in a model that included both frequency and length as covariates. Morever, our sample size is bigger.}.
Despite this cross-linguistic variation, both phonological and semantic
connectivity remain significant predictors in the global model.

The present study provided a comprehensive analysis of how lexical
connectivity influences the age of acquisition of nouns in toddlers. We
compared two network growth scenarios and assessed the relative
contribution of phonological and semantic information in 8 languages.
Part of the findings largely replicate the results obtained in Hills et
al. (2009), i.e., semantic networks (based on free associations) grow by
preferential acquisition, not by preferential attachment. Another
finding was that phonological networks also grow primarily by
preferential acquisition, especially when both scenarios (PAT and PAC)
were pitted against each other in the same model. These findings
generalize well across languages. Moreoever, both semantic and
phonological connectivity in the learning environment (i.e., PAC)
predict growth in a consistent way across many languages. However, when
pitted against other known predictors of age of acquisition (word
frequency and length), the effect of word connectivity shows a
cross-linguistic variation, predicting part of learning in some
languages, but not in others. Despite this coross-lingusitic variation,
both phonological and semantic connectivity contribute to the overall
learning (when data is pooled across languages).

>>>>

\title{Word Learning as Network Growth: a Cross-linguistic Analysis}

@@

\begin{abstract}
Children tend to learn first the words that are highly connected in the
lexical network (i.e., words that are connected to a variety of other
words). The process by which connectivity in both the phonological and
semantic netwoks come to influence word learning is still not very well
understood. One view suggests that learning is driven primarily by
\emph{internal} connectivity in the child's available lexicon. Another
view suggests that \emph{external} connectivity in the learning
environment is what makes words salient and readily learned. The present
study tests both scenarios systematically in both the phonological and
the the semantic domains, and across 8 languages. We generate the
results obtained by Hills et al. (2009) with English semantic networks,
and show that external connectivity in the learning environment, rather
than internal connectivity, drives growth of both semantic and
phonological networks. Finally, we demonstrate that these effects hold
cross-linguistically.

\textbf{Keywords:}
semantic network, phonological network, word learning
\end{abstract}

@@

What factors shape vocabulary learning over the course of early
childhood? To investigate this question, scientists have adopted
multiple research strategies, from controlled laboratory experiments to
corpus analysis. One strategy consists in documenting the timeline of
words' acquisition, and studying the properies that make words easy or
hard to learn. For example, within a lexical category, words that are
more frequent in child-directed speech are acquired earlier (J. C.
Goodman, Dale, \& Li, 2008). Other factors include word length, the mean
length of utterances in which the word occurs, and how concrete the word
is (see Braginsky, Yurovsky, Marchman, \& Frank, 2016).

Besides these word-level properties, researchers found that the lexical
structure (that is, how words relate to each other) also influences the
age of acquisition of words. The lexical structure is best characterized
in terms of a network where each node represents a word in the
vocabulary, and each link between two nodes represents a relationship
between the corresponding pair of words. Previous studies investigated
early vocabulary networks based on different word relations such as
shared semantic features, target-cue relationship in free association
norms, co-occurrence in child directed speech, and phonological
similarity. These studies found that children tend to produce earlier
the words that have higher neighborhood density (i.g., high connectivity
in the network) both at the phonological and the semantic level
(Beckage, Smith, \& Hills, 2011; Engelthaler \& Hills, 2017; Hills,
Maouene, Riordan, \& Smith, 2010; Hills, Maouene, Maouene, Sheya, \&
Smith, 2009; Stella, Beckage, \& Brede, 2017; Stokes, 2010; Storkel,
2009).

A crucial question that naturally emerges from these analyses is how
does network connectivity influence word learning? Steyvers \& Tenenbaum
(2005) suggested that the effect of connectivity is the consequence of
how the lexical network gets constructed in the child's mind. According
to this explanation, known as Preferential Attachment (PAT), highly
connected words in the child's lexicon tend to ``attract'' more words
over time, in a rich-get-richer scenario (Barabasi \& Albert, 1999). In
other words, what predicts word learning is \emph{internal} connectivity
in the child's early lexicon. In contrast, Hills et al. (2009) found
that what biases the learning is not the connectivity in the child's
internal lexicon but, rather, \emph{external} connectivity in the
learning environment. They called this alternative explanation
Preferential Acquisition (PAC). Figure \ref{fig:growth} shows an
illustration of both growth scenarios with the same simplified network.

\caption{\label{fig:growth}Illustration of the two growth models with the same network. Filled circles (I1-I4) represent known words (internal lexicon), and empty circles (E1 and E2) represent words that have not been learned yet (external lexicon). Black lines represent links that are relevant in each growth scenario, and gray lines represent links that are irrelevant. For PAT, each candidate node is characterized with the average degree (i.e., number of links) of the existing nodes that it would attach to. Thus, according to PAT, the node E1 is more likely to enter the lexicon first. For PAC, each candidate node is characterized with its degree in the entire network. According th PAC, the node E2 is more likely to enter the lexicon first.}\label{fig:growth}
\end{figure}
\end{CodeChunk}

Studies that investigted lexical network growth focused on semantic
networks using English data only (Hills et al., 2010, 2009; Steyvers \&
Tenenbaum, 2005). The novelty of the current study is that it
investigates whether phonological networks, like semantic networks, grow
by PAC, or if they rather grow by PAT. It also provides a systematic
comparision of network growth scenarios in both the phonological and the
semantic domains and assesses their relative contribution to the
learning process. Moreover, it tests the generality of the findings
across eight languages.

The paper is organized as follows. First, we describe the dataset we
used, the procedure we followed to construct both semantic and
phonological networks. Next, we show the results of the various analyses
we conducted comparing network growth scenarios across languages. Then
we test the overall contribution of the networks in the learning process
relative to other known predictors of age of acquisition (frequency and
length). Finally we discuss the major results in relation to known
experimental facts in language development.

@@

We used data from Wordbank (Frank, Braginsky, Yurovsky, \& Marchman,
2017), an open repository aggregating cross-linguistic language
developmental data of the MacArthur-Bates Communicative Development
Inventory (CDI), a parent report vocabulary checklist. We used the
\emph{Words and Sentence} version of the CDI which contains the
productive vocabulary of toddlers (age varied between 16 to 36 months).
Following previous studies (Hills et al., 2009; e.g., Storkel, 2009), we
restricted our analysis to the category of nouns, since there are known
discrepancies between the learning of nouns, predicates, and function
words (e.g., Braginsky et al., 2016). The age of acquisition was defined
by the month at which a word was produced by at least 50\% of children
(J. C. Goodman et al., 2008).

We obtained these nouns in eight languages: Croatian, Danish, English,
Italian, Norwegian, Russian, Spanish, and Turkish. We used the subset of
nouns that had entries in the Florida Association Norms (see below).
Since these norms are available only in English, we used the translation
of non-English words provided by Braginsky et al. (2016). This allowed
us to use the English association norms across languages. Table
\ref{tab:stats} gives an overview of the data used. The translation of
non-English words is still an ongoing process. Note, however, that all
languages have at least 60\% of nouns translated.

\begin{table}[H]
\centering
\begin{tabular}{rlrrr}
  \hline
 & language & total & translated & normed \\ 
  \hline
1 & Croatian & 253 & 177 & 170 \\ 
  2 & Danish & 295 & 198 & 187 \\ 
  3 & English & 296 & 296 & 274 \\ 
  4 & Italian & 311 & 203 & 194 \\ 
  5 & Norwegian & 305 & 193 & 186 \\ 
  6 & Russian & 311 & 311 & 285 \\ 
  7 & Spanish & 240 & 173 & 163 \\ 
  8 & Turkish & 293 & 175 & 164 \\ 
   \hline
\end{tabular}
\caption{\label{tab:stats}Total number of productive nouns in the CDI (left). We used a subset of these nouns that had available English translations (middle). The final set consisted of nouns that had both available translations as well entries in the Free Association Norms (right).} 
\end{table}

@@

We used as an index of semantic relatedness the Florida Free Association
Norms (Nelson, McEvoy, \& Schreiber, 1998). This dataset was collected
by giving adult participants a word (the cue), and asking them to write
the first word that comes to mind (the target). For example, when given
the word ``ball'', they might answer with the word ``game''. A pair of
nodes were connected by a directed link from the cue to the target if
there was a cue-target relationship between these nodes in the
association norms. Each node was characterized by its ``indegree'',
which represents the number of links for which the word was the target.
Following Hills et al. (2009), we used as a proxy for the learning
environment, the network constructed using the full set of nouns in the
CDI data in each language (e.g., in English this corresponds to the
network by 30 months).

We used as a measure of phonological relatedness between two nodes the
Levenshtein (edit) distance between their phonological forms. The
measure counts the minimum number of operations (insertions, deletions,
substitutions) required to change one string into another. We generated
approximate International Phonetic Alphabet (IPA) trnascriptions from
the orthographic transcription, across languages, using the open source
text-to-speech software
\textbf{\href{http://http://espeak.sourceforge.net/}{Espeak}.}

In previous studies, two nodes were linked if they had an edit distance
of 1 (Stokes, 2010; Storkel, 2009). However, in these previous studies
the network was built using an adult vocabulary. Here, similar to the
approximation we made for the semantic network, we used the full set of
nouns in the CDI data as a proxy for phonological connectivity in the
learning environment. However, since the children's vocabulary contains
very few word pairs with an edit distance of 1, the resulting network
was too sparse and uninformative. Thus, we increased the threshold from
1 to 2, that is, two node were related if their edit distance was equal
to 1 or 2. Each node was characterized with its degree, i.e., the number
of links it shares with other words.

We start by examining how connectivity in the global netowrk influences
the age of acqusition. Regardless of the growth mechanism that might
have given rise to this netowrk (i.e., PAC or PAT), we expect nodes with
the higher degrees in this network to be learned earlier than nodes with
the lower degrees. The network was constructed using nouns at the oldest
age for which we have CDI data. Figure \ref{fig:data_all} shows how the
age of acquisition for each word varies as a function of its degree (or
indegree for the semantic network). For ease of visual comparison, the
predictor (i.e., the degree) was centered and scaled across languages.
The plots show, overall, a negative correlation between the month of
acquisition and the degree, indicating that nouns with higher degrees
are generally learned earlier.

@@

\caption{\label{fig:pred_ind}Evaluation of network growth scenarios both individually (dotted), and when combined in the same growth model (solid). Each dot represents the mean of the posterior distribution of the corresponding growth parameter, with ranges representing 95\% credible intervals (computed using the highest density intervals).}\label{fig:pred_ind_img}
\end{figure*}
\end{CodeChunk}

@@

\subsubsection{How does each growth scenario predict noun
development?}\label{how-does-each-growth-scenario-predict-noun-development}

To test the network growth scenarios, we fit different growth models to
the data. We proceeded as follows. We calculated the probability that a
word \(w_i\), with a growth value \(d_i\) would enter the lexicon at a
given month, using a softmax function:

@@

Where \(\beta\) is the fitting parameter. A positive value of \(\beta\)
means that words with higher growth values \(d_i\) are acquired first,
and a negative value means that words with lower growth values are
acquired first. As explained in Figure \ref{fig:growth}, each candidate
word \(w_i\) is characterized with a growth values \(d_i\) that depends
on the growth scenario tested. The normalization includes all words that
could be learned at that month.

We estimated the parameter \(\beta\) using a Bayesian approach. The
inference was performed using the probabilisitic programming language
WebPPL (N. Goodman \& Stuhlmüller, 2014). We started with a uniform
distribution, and at each month, we computed the likelihood function
over words that could possibly enter the lexicon at that month, fit to
the words that have been actually learned at that month (using formula
1). We ended up with a posterior distribution of \(\beta\), which we
summarized in Figure \ref{fig:pred_ind}.

Besides fitting a growth model to the data, we conducted another
evaluation. This second evaluation consists in determining, in each
month, the model-dependent growth value distribution of all words that
could possibly be learned at this month, and then computing the z-score
of each learned word with respect to this distribution. For each model,
we tested if the distribution constituted by the z-scores of all learned
words was different from zero, using a one-sample t-test.

The results from both evaluations were very similar and lead essentially
to the same
conclusions\footnote{we do not show here the resuts of the second evaluation because they were redundant with the results of the first evaluation}.
For the semantic network, the results replicate Hills et al.'s finding
in English, which was that the semantic network grows by PAC, not by
PAT. Moreover, it generalizes this finding to all other languages (with
the exception of Russian). For the phonological network, the results
show that, generally speaking, PAC again fits the developmental
trajectory better than PAT. We note however that PAT, though weaker,
fares better for the phonological networks (where it predicts part of
the growth process in some languages such as Croatian, English,
Norwegian and Russian) than it does for the semantic networks (where it
is rather universally unpredictive).

\subsubsection{What is the relative contribution of each growth
model?}\label{what-is-the-relative-contribution-of-each-growth-model}

Above we evaluated the network growth scenarios individually. As a next
step, we analysed their relative contribution to the learning process.
This was done through adding more fitting parameters to the model, that
is, by substituting \(\beta d_i\) in formula (1) with:
\[\beta_{1} d_{i, 1} + \beta_{2} d_{i, 2} + \beta_{3} d_{i, 3} + \beta_{4} d_{i, 4}\]
where the indices represent the 4 networks: semPAT, semPAC, phonoPAT and
PhonoPAC. Using the same fitting technique, we obtained the values shown
in Figure \ref{fig:pred_ind}. In term of growth scenarios, we found that
PAC dominates the learning. In particular, though we found previously
that phonological PAT predicted part of the learning in some languages
when tested individually, here PAT appears to lose its predictive power
when pitted against phonological PAC. In terms of domain, both
phonological and semantic networks appear to contribute to learning,
although the phonological network appears to be stronger and more
reliable across languages. In summary, the findings show that both
semantic and phonological networks grow primarily by PAC, and that,
generally speaking, semantic and phonological networks both contribute
to the learning process.

@@

\subsection{Comparison to other known predictors of age of
acquisition}\label{comparison-to-other-known-predictors-of-age-of-acquisition}

We saw that the way semantic and phonological information is structured
in the learning environment (i.e., PAC) contribute to noun learning
across languages. However, we know that other factors influence learning
as well (e.g., Braginsky et al., 2016). In what follows, we will
investigate how semantic and phonological connectivity interact with two
other factors. The first one is word frequency, a well studied factor
shown to predict the age of acquisition in a reliable fashion (e.g. J.
C. Goodman et al., 2008). The second factor is word length, which
correlates with phonological connectivity (Pisoni, Nusbaum, Luce, \&
Slowiaczek, 1985).

Since PAT was uninformative, we dropped it from this analysis. Thus, we
no longer needed to fit the growth model month-by-month as in the
previous section. In fact, word utilities in the case of PAC are fixed,
they do not depend on previously learned words. A more direct way to
assess and compare the contribution of PAC in relation to other factors
is through conducting linear regressions, where connectivity in the
learning environment (at both the phonological and semantic level),
frequency and length predict the age of acquisition.

We used the frequency estimates from Braginsky et al. (2016) where
unigram counts were derived based on CHILDES corpora in each language.
For each word, counts included words that shared the same stem (e.g.,
cats counts as cat), or words that were synonymous (e.g.~father counts
as daddy). For word length, we used our generated IPA transcription.

We conducted two analyses. We fit a linear regression for each language,
and we fit a linear mixed-effect to all the data pooled across
languages, with language as a random effect. Figure
\ref{fig:regressions_img} shows the coefficient estimate for each
predictor in each language, and figure \ref{fig:regressions_all_img}
shows the coefficient estimates for all languages combined (all
predictors were centered and scaled). The findings were as follows.
Overall, frequency is the largest and most consistent predictor of age
of acquisition. Word length predicts learning in some languages such as
Croatian and Norwegian, but not in others (including English). It
remains, however, a significant predictor in the global model. As for
the factors of interest, i.e., semantic and phonological connectivity,
we also found cross-linguistic differences. The phonological
connectivity contributes to learning in languages such as Croatian,
English and Russian, whereas semantic connectivity contributes to
learning in Turkish, Spanish and to some extent in Croatian, but
interestingly not in
English\footnote{Semantic connectivity does not explain variance in English data beyond that explained by phonological connectivity, frequency and length. This contrasts with the original finding in Hills et al. 2009. However, in this previous study, semantic connectivity was not tested in a model that included both frequency and length as covariates. Another important difference is the number of words tested. Our study uses a larger set of nouns.}.
Despite this cross-linguistic variation, both phonological and semantic
connectivity remain significant predictors in the global model.

The present study provided a comprehensive analysis of how lexical
connectivity influences the age of acquisition of nouns in toddlers. We
compared two network growth scenarios and assessed the relative
contribution of phonological and semantic information in 8 languages.
Part of the findings largely replicate the results obtained in Hills et
al. (2009), i.e., semantic networks (based on free associations) grow by
preferential acquisition, not by preferential attachment. Another
finding was that phonological networks also grow primarily by
preferential acquisition, especially when both scenarios (PAT and PAC)
were pitted against each other in the same model. These findings
generalize well across languages. Moreoever, both semantic and
phonological connectivity in the learning environment (i.e., PAC)
predict growth in a consistent way across many languages. However, when
pitted against other known predictors of age of acquisition (word
frequency and length), the effect of word connectivity shows a
cross-linguistic variation, predicting learning in some languages, but
not in others. Despite this cross-lingusitic variation, both
phonological and semantic connectivity contribute to the overall
learning (when data is pooled across languages).

@@

\caption{\label{fig:regressions_all_img}Estimates of predictor coefficients in the combined model pooling across languages . Ranges indicate 95\% confidence intervals. Fade color indicates estimates of PAC predictors in a model that does not include frequency and length as covariates.}\label{fig:regressions_all_img}
\end{figure}
\end{CodeChunk}

A major results of the study is that children start by learning words
that have high phonological and semantic similarity with a variety of
other words in the learning environment, not in the child's available
lexicon. This suggests that children are sensitive to connectivity even
without having first acquired the connected words. How can children
indirectly detect highly connected words, and why would such words be
more readily learned? In the semantic case, free association can be
predicted through the patterns of word co-occurrence (Griffiths,
Steyvers, \& Tenenbaum, 2007), meaning that highly connected words tend
to be the words that co-occur with many other words in various contexts.
One possibility, suggested by Hills et al. (2010), is that the referents
of such words are more easily disambiguated from other potential
referents. Such effect was demonstrated by cross-situational learning
experiments (Smith \& Yu, 2008). In the phonological case, connectivity
is inherently correlated with phonotactic probability (Vitevitch, Luce,
Pisoni, \& Auer, 1999). That is, highly connected words tend to be made
of frequent sound sequences. We know even infant (whose vocabulary is
still very rudimentary) develop a sensitivity for high frequency sound
sequences in the ambient language (Jusczyk, Luce, \& Charles-Luce,
1994). Interestingly, it was shown that phonotactic probability
facilitates learning and recognition of novel words in toddlers and
preschoolers (MacRoy-Higgins, Shafer, Schwartz, \& Marton, 2014;
Storkel, 2001). Thus, children's ability to keep track of co-occurrence
statistics (both at the semantic and phonological level) might explain
their sensitivity and preference for highly connected words in the
learning envirnoment.

Finally, this study shares a number of limitations with previous studies
using similar datasets. For instance, we used normative age of
acquisition. However, individual children may follow different learning
trajectories. The use of longitudinal data would allow us to assess how
the aggregate behavior relates to these individual trajectories (e.g.,
typical vs.~late talkers, Beckage et al., 2011). Moreover, the results
obtained in this study provide correlational but not causal evidence.
Thus, conclusions drawn from our network analysis would require parallel
evidence, especially from controlled behavioral experiments.

\vspace{1em}

\fbox{\parbox[b][][c]{7.3cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/afourtassi/networks}}}
\vspace{1em}
